**Suggestions for improving invited speaker diversity to reflect trainee diversity**

### 2020-02-07 submission/manuscript.Rmd
finish revisions and format resubmission

---
title: '**Suggestions to improve invited speaker diversity and reflect trainee diversity**'
csl: mbio.csl
fontsize: 10pt
output:
  pdf_document:
    includes:
      in_header: header.tex
    keep_tex: yes
geometry: margin=1.0in
bibliography: references.bib
---

@@

Running title: Suggestions to improve invited speaker diversity

@@

Ada K. Hagan, Ph.D.$^{1,2\dagger}$, Rebecca M. Pollet, Ph.D.${^2}$, and Josie Libertucci, Ph.D.${^3}$

@@

1\. Alliance SciComm & Consulting, Linden, Michigan

2\. Department of Microbiology & Immunology, University of Michigan, Ann Arbor, Michigan

3\. Farncombe Family Digestive Health Research Institute, Department of Medicine, McMaster University, Hamilton, Ontario, Canada

@@

The biomedical sciences have a problem retaining women and underrepresented minorities in academia. Despite increases in the representation of these groups in faculty candidate pools, they are still underrepresented at the faculty level, particularly at the Full Professor level. The lack of diverse individuals at the Full Professor level contributes to the attrition of women and under-represented minorities, as it confirms unconscious biases. The presence of unconscious biases contribute to feelings of not belonging by trainees and are amplified by visual representation of who is presented as the “top scientist in their field”. Top scientists are not only defined by the attainment of Full Professorships, but also through invited seminar series. Invitations for faculty to present their research at other university departments is highly valued offer that provides an opportunity for collaborations and networking. However, if invited speakers do not represent the demographics of current trainees, these visual representations of successful scientists may contribute to decreased attitudes of self-identification as a scientist, ultimately resulting in trainees leaving the field or the academy. In this study, we compare invited-speaker demographics to the current trainee demographics in one microbiology and immunology department and find that trainees are not proportionally represented by speakers invited to the department. Our investigation prompted changes in policy for how invited speakers are selected in the future to invite a more diverse group of scientists. To facilitate this process, we developed a set of tips and a web-based resource that allows scientists, committees, and moderators to identify members of under-served groups. These resources can be easily adapted by other fields or sub-fields to promote inclusion and diversity at seminar series’, conferences, and colloquia.

@@

Long-standing systemic bias, sexism, and racism have contributed to the under-representation of many racial and ethnic groups, as well as women, in science, technology, engineering, and math (STEM) fields [@Martinez2018; @AllenRamdial2014; @Fang2000; @Gibbs2014]. Specifically, within the field of biomedical research in the United States, the proportion of underrepresented minorities at the full professor level has remained consistently low at 4% (survey data taken from the NIH from 2001 to 2013), compared to the U.S. population, which is 32.3% [@Meyers2018;@nsf_2014]. Similar discrepancies exist for women in biomedical sciences as full professorships are currently held mostly by men [@Jena2015; @Rotbart2012].  As demographics of faculty within the biomedical sciences remains skewed towards Caucasian men, the demographics of trainees (graduate students and postdocs) are becoming more diverse [@Meyers2018].

Policy changes are needed to support inclusion of all individuals, particularly in the biomedical sciences[@Coe2019]. To increase retention of historically under-represented minorities (HURM), non-Caucasian/non-HURM (NCNH) individuals, and Caucasian women in biomedical fields, it is important for trainees to have visual representations of themselves as scientists. The importance of representation in retaining a diverse group of individuals in STEM fields is supported by social role theory [@Eagly1984]. Individuals make inferences about characteristics that are needed to be successful in a given role by examining individuals that most occupy that role [@Eagly1984; @Carter2018]. However, there is a lack of diverse scientific experts in academia so underrepresented minorities are not seeing adequate visual representations of themselves at the faculty level. Therefore, trainees who do not see representation of themselves in senior faculty positions, may decide that they do not possess the characteristics that are required to succeed.

Invited seminar series are common within biomedical departments across the United States [@nittrouer_gender_2018]. Usually, seminar series' consist of faculty members selecting a scientist from another institution to visit their university and present their research, as well as meet with other faculty members and trainees. Named lectureships follow the same format but are decided by committee and are considered more prestigious because they are named in honor of prominent local scientists. These seminar series and lectureships provide an opportunity for trainees to be exposed to research outside of their department. Additionally, being an invited speaker provides the scientist with an opportunity to make future collaborations and build their own _curriculum vitae_ (CV). Scientists invited to give seminars are widely regarded as successful and the top in their field. Thus, if trainees are constantly being exposed to “the top scientist in their field”, according to social role theory, it signals who is successful in that field. While some have examined this issue by studying and promoting the inclusion of more women speakers at conferences, how department speaker series compare to the trainee diversity of that department is unknown [@kalejta_gender_2017; @casadevall_presence_2014; @klein_speaking_2017].

In this study, we examine and compare the proportion of HURM, NCNH, and women invited speakers to Caucasian men in the Department of Microbiology and Immunology at the University of Michigan. Additionally, we compare invited-speaker demographics to the current trainee demographics as a means to gauge if trainee demographics are being represented accordingly throughout the seminar series. Following our investigation, we proposed a policy change to the Department of Microbiology and Immunology in how invited speakers are selected as a means to promote inclusion in our department and reduce unconscious bias. In order to facilitate inviting a more diverse group of scientists, we developed a set of resources that allow scientists, within the fields of microbiology and immunology, to self-identify as having an under-represented or under-served identity including: HURM, non-Caucasian/non-HURM, or a Caucasian woman. These resources will promote inclusion and diversity by providing greater representation of all scientists and will provide hosts an opportunity to invite a more diverse group of scientists.

## Methods
Each academic year, each faculty member in the Department of Microbiology and Immunology at the University of Michigan has the opportunity to invite one speaker per year for a weekly seminar series. Some of these seminar slots are dedicated to named lectureships, which are decided by committee, and three trainee-invited speakers. We analyzed the demographics of invited speakers and faculty hosts for five academic years (Fall 2014 - Spring 2019), and compared them to the current trainees when the data were analyzed (Spring 2019). Each speaker was only counted once and those listed as departmental faculty members or as a “host” at any point could not also be considered “invited speakers”. The list of faculty hosts was used as a proxy for faculty demographics since as hosts, these faculty members are visible representatives of the department. There were a total of {`r`} invited speakers and paired hosts. The trainees were identified using departmental email lists that included masters students, doctoral students, and post-doctoral fellows. There were {`r`} students and {`r`} post-doctoral fellows.

This is a retrospective study, thus speakers were not asked for their identities at the time of visit. Instead we hand-coded proxy demographics of the speakers, faculty hosts and trainees using personal knowledge and publicly available photos, and CVs [@Broderick2019; @Kimery2011; @Helmer2017; @Fox2019; @Gilbert1994]. Information from CVs, such as undergraduate institutions (e.g., Puerto Rico vs mainland U.S.) and activity in HURM groups (e.g., SACNAS) helped inform our demographics. Because these data were collected from publicly available sources, this study was not submitted to an IRB for consideration. The presenting gender of each individual was assigned using a binary system (man/woman). Due to the low number of individuals in the study, race/ethnicity demographics were split in three groups: Caucasian, Historically Under-represented Minority (HURM), and Non-Caucasian/Non-HURM (NCNH), each with a binary (yes/no) possibility. Caucasian was assigned using the current U.S. Census definition where those of Middle Eastern, European, and Russian descent are included. URM individuals include those of African-American, Indigenous American, Alaskan/Hawaiian Native, Latinx and/or Hispanic heritage (20 U.S. Code § 1067k), we use the HURM designation to recognize the history of enslavement and active oppression in the U.S.. The NCNH group predominantly included Asian/Asian-Americans, but also African immigrants [@Okonofua2013].

@@

trainee_summary <- tidy_trainee %>% 
  select(-Gender) %>% 
  filter(value == "y") %>% distinct_() %>% 
  group_by(demographic) %>% summarise(n = n()) %>% 
  mutate(prop = get_percent(n, sum(.$n)))
```
To understand the representation of women, we compared the proportion of women in each academic role. At the trainee level, more than half of students and postdoctoral fellows were women. That dropped to `r gender_combined[[1,4]]`% of faculty hosts and `r gender_combined[[3,4]]`% of the invited speakers (Fig. 1A). Of `r sum(lect_summary$n)` lectureships over the five year period, `r lect_summary[4,4]`% were awarded to women.

@@

This study found that the proportion of HURM and NCNH invited speakers were under-representative of the trainee populations for each group. Additionally, within the last 5 years, no HURM or NCNH woman was awarded a lectureship. This means that the department is not providing non-Caucasian trainees with adequate representation of successful scientists and failing to support an inclusive environment in terms of visual faculty representation. We also found that the proportion of women as faculty hosts and speakers in our study population is equivalent to global estimates that 40% of microbiologists are women, though women only represent about 30% of academic biomedical faculty [@allagnat_gender_2017;@Jena2015]. Women are also over-represented as graduate students and postdoctoral fellows in this department. Overall, Caucasian scientists are over-represented as host faculty and invited speakers, compared to their presence as trainees, particularly when lectureships were considered.

We have not been able to identify any publications examining scientific speaker diversity beyond gender [@nittrouer_gender_2018]. This seems to be the first, which is concerning since conclusions drawn from gender-based studies are often framed, and considered, to be applicable to other marginalized groups (e.g., HURM). This is a flawed assumption. While there is no doubt some overlap, each group remains marginalized due to a unique complex set of factors that cannot always be solved by gender-based solutions. The historical exclusion of HURMs by U.S. institutions means that they have a particular responsibility to improving the academic experience for these populations [@harvey_diversity_2011]. We therefore call on  U.S. institutions to apply this framing to their discussions and research.

Departments have different processes and criteria for selecting invited speakers, but it is a matter of pride to bring the best scientists possible. The barriers to achieving faculty status for HURMs, Caucasian women, and NCNH may also impact their speaking invitations. For instance, the perceived prioritization and commitments of women to family over work may cause faculty to doubt their acceptance of a speaking invitation, despite the prestigious nature of these invitations and evidence that men and women accept at similar rates [@nittrouer_gender_2018; @xu_gender_2008]. As a result, the faculty member may invite a different colleague who they feel is more likely to agree (and is a man). It may also be that the definition of “best” poses a problem to under-represented and under-served groups (e.g., women, HURM, and Asian) who are held to stricter competency standards [@BlairLoy2017;@noauthor_seeking_2013]. Some departments may only invite tenured faculty, which severely limits the number of potential speakers who are Caucasian women or non-Caucasian. Yet, another scenario is that pre-tenure faculty members invite prestigious, tenured faculty in their field to network and secure letters for their own tenure package. The increased burden of women and non-Caucasian scientists to prove competency decreases their likelihood to be considered for either tenure or as possible source of tenure letters. In particular, the proportion of HURM faculty at the Assistant and Associate Professor level is currently higher than at Full Professor so it will be difficult to increase speaker diversity if early-career researchers are not being considered [@whittaker_retention_2015]. Even when HURM speaker rates match the proportion of HURM faculty employment, HURM trainees will be represented at a significantly higher proportion. We argue that inclusion of marginalized faculty in seminar series is an important factor to increasing their representation among Associate and Full Professors. It is just one aspect of larger institutional change that is needed, but one that will benefit trainee experiences and the CVs of marginalized faculty [@johnson_msphere_2019].

We recognize that our proxy demographics are a limitation of the analysis and want to acknowledge that biological sex (male/female) is not always equivalent to the gender that an individual presents as (man/woman), which is also distinct from the gender(s) that an individual self-identifies as. We also want to acknowledge that there are many other identities that are not captured in this limited analysis. While our pilot study combined 5 years worth of seminars, our n is still quite low and we did not have other departments for comparison. Consequently, our results can not be generalized to other departments, fields, or universities. Another limitation to looking at a single department is that trainees may interact with faculty from many other departments, depending on their research and interests. Therefore, the individual experience of representation would vary by trainee. Future research needs to consider multiple departments, universities, and/or fields to bolster generalizability. There is a paucity of research on speaker identities other than their gender, so this also needs to be addressed in future studies. However, we caution that representation is a shallow metric on which a single-minded focus can cause more harm than help [@Iverson2007]. We recommend that future research also survey trainee and speaker experiences and trainee participation in seminar series to better understand the dynamics at play.

@@

We also used this opportunity to begin a conversation about the purpose of seminar speakers. Seminar speakers are sometimes invited to highlight their latest high-impact paper or to share the arc of discoveries they have made over several years of their career. While both are certainly worthwhile, there are other benefits to be gained from interacting with seminar speakers such as how to apply new techniques and how research is framed outside research-focused universities (Table 1). Thinking more broadly about what material is valuable during a seminar series may lead to more speakers from under-represented and under-served backgrounds, as well as more diversity of career paths. For some institutions, these suggestions represent more of a structural change to the departmental seminar series as speakers focusing on techniques or from non-research-intensive universities are using invited as part of professional development series. If these changes are to be implemented, many members of the department must agree to the value of including these seminars in the main departmental seminar series and these expectations must be clearly communicated by the leadership (Table 1). Departmental leadership can also ask individuals and groups to consider their own unconscious biases that may be impacting their speaker nominee list to compact some of the barriers to inviting diverse speakers.
Presented with these ideas, several members of the Department of Microbiology and Immunology at the University of Michigan expressed interest in specific resources they might use to identify individuals from diverse identities, careers, and institutions. One suggestion is to use resources that are being organized by professional societies such as the American Society for Microbiology (ASM) and the Society for Advancing Chicanos/Hispanics & Native Americans in Science (SACNAS) (Table 1). We also chose to develop “Diversify” resources for the microbiology and immunology fields that provide a list of scientists from under-represented and under-served groups, that is not associated with a specific professional society. More information on the type of resources and how to establish a Diversify list is below. Using list resources like those available from professional societies as well as our “Diversify” resources are particularly useful as social science research has shown that the human brain is much better at recognizing and using information (such as a strong scientific speaker) from a list than it is at recalling the same information from memory [@Loftus1971; @Johnson2014].

We caution, however, that it is not enough to invite diverse speakers. An inclusive environment must be built within the department. Start by inviting all speakers to spend a few minutes describing their personal science journey and providing time for trainees to engage with the speaker. Trainee-speaker interactions can be encouraged by ensuring that trainees are the first to ask questions at the seminar’s conclusion and by scheduling a dedicated meeting time for trainees. As speaker schedules are being designed, departments should consider how they can foster an inclusive atmosphere during the speaker visits. Speakers should be asked prior to their visit if they have any dietary, movement, or other restrictions that should be accommodated during the visit. The identities of individuals the speaker is meeting with during their visit may also need to be considered; this is not to say that all identities of a speaker should be matched on their schedule or that a HURM speaker must have a meeting a HURM faculty member, but take care to ensure that speakers are meeting with faculty that reflect the diversity of thought and identity in the department. If a portion of the department is consistently not represented on speakers’ schedules (or extending invitations) this may reflect an opportunity for increased inclusivity in the department. Finally, speakers should be provided with ample opportunity to request meetings, not only with faculty but also with student groups or campus administrators who share similar interests. Through these steps departments can increase the diversity of speakers invited to their seminars while also increasing the impact of the seminar speakers.

@@

The website provides an interface to the Google forms and spreadsheets with template pages for viewing the list, adding a name to the list, and finding additional resources. Importantly, our website creation tool is hosted for free by GitHub, which provides a free website for each GitHub organization. Basic tools and skills required to set up a Diversify site include knowledge of, or experience with, the version control tool git, the web-tool GitHub, and a text editor. A tutorial in the DiversifyMicrobiology repository on GitHub provides links to these resources and instructions for adapting the tool to your own field [@Hagan2019_micro]. We caution creators of Diversify lists that the data voluntarily submitted to these lists is not eligible for study. IRB approval must be obtained prior to launching the list if that is a goal.

@@

To increase the retention of Caucasian women, HURM and NCNH trainees in the biomedical sciences, they must also be represented as experts. However, the invited speaker diversity at one department does not represent the diversity of trainees. To facilitate the identification and recruitment of individuals in these historically under-served groups, we have built a tool to create self-nominated, field-specific lists.

@@

## Author Contributions
A.K.H. collected the data, assigned demographics, analyzed the data, created the website, and wrote the methods and results. R.M.P. created the Google lists, forms, and website content and the description of their maintenance. J.L. wrote the introduction and provided conceptual advice. A.K.H. and J.L facilitated the policy change to the UM Department of Microbiology and Immunology. All authors contributed to preparing the final manuscript.

@@

![](Table_1.pdf)

![**The demographics of invited speakers, hosting faculty, and trainees.** A) The proportion of women in each academic role. B) The proportion of each academic role represented by individuals that are Caucasian (left), Historically Underrepresented Minorities (HURM, center) or Non-Caucasian/Non-HURM (NCNH, right). C-D)The percent of lectureships awarded to individuals that are C) Caucasian, HURM, or NCNH and D) Caucasian, HURM, or NCNH by gender.](Figure_1.jpg)

@@

## References

>>>>

---
title: '**Suggestions for improving invited speaker diversity to reflect trainee diversity**'
csl: mbio.csl
fontsize: 10pt
output:
  word_document: default
  pdf_document:
    includes:
      in_header: header.tex
    keep_tex: yes
  html_document:
    df_print: paged
geometry: margin=1.0in
bibliography: references.bib
---

@@

Running title: Suggestions for improving invited speaker diversity

@@

Ada K. Hagan, Ph.D.$^{*1\dagger}$, Rebecca M. Pollet, Ph.D.${^1}$, and Josie Libertucci, Ph.D.$^{**1}$

@@

1\. Department of Microbiology & Immunology, University of Michigan, Ann Arbor, Michigan

*Current address: Alliance SciComm & Consulting, Linden, Michigan

**Current address: Farncombe Family Digestive Health Research Institute, Department of Medicine, McMaster University, Hamilton, Ontario, Canada

@@

Within the field of biomedical research in the United States, the proportion of underrepresented minorities at the full professor level has remained consistently low, even though trainee demographics are becoming more diverse.  Underrepresented groups face a complex set of barriers to achieving faculty status including imposter syndrome, increased performance expectations, and patterns of exclusion. Institutionalized racism and sexism have contributed to these barriers and perpetuated policy that excludes underrepresented minorities. These barriers can contribute to decreased feelings of belonging, which may result in decreased retention of underrepresented minorities. Though some universities have altered their hiring practices to increase the number of underrepresented minorities in the applicant pool, these changes have not been sufficient. Here we argue that departmental invited seminar series can be used to provide trainees with scientific role-models and increase their sense of belonging, while institutions work towards more inclusive policy. In this study, we investigated the demographics (gender and race) of invited seminar speakers over 5 years to the Department of Microbiology and Immunology at the University of Michigan. We also investigated current trainee demographics and compared them to invited speaker demographics to gauge if our trainees are being provided with representation of themselves. We found that invited speaker demographics were skewed towards Caucasian men, and our trainee demographics were not being  represented. From these findings, we proposed policy change within the department to address how speakers are being invited with the goal of increasing speaker diversity to better reflect trainee diversity. To facilitate this process, we developed a set of suggestions and a web-based resource that allows scientists, committees, and moderators to identify members of under-served groups. These resources can be easily adapted by other fields or sub-fields to promote inclusion and diversity at seminar series’, conferences, and colloquia.

@@

Long-standing systemic bias, sexism, and racism have contributed to the under-representation of many racial and ethnic groups, as well as women, in science, technology, engineering, and math (STEM) fields [@Martinez2018; @AllenRamdial2014; @Fang2000; @Gibbs2014]. Specifically, within the field of biomedical research in the United States, the proportion of underrepresented minorities at the full professor level has remained consistently low at 4% (survey data taken from the NIH from 2001 to 2013) [@Meyers2018;@nsf_2014]. Similar discrepancies exist for women in biomedical sciences as full professorships are currently held mostly by men [@Jena2015; @Rotbart2012].  As demographics of faculty within the biomedical sciences remain skewed towards Caucasian men, the demographics of trainees (graduate students and postdocs) are becoming more diverse [@Meyers2018].

Policy changes are needed to support inclusion of all individuals, particularly in the biomedical sciences, since underrepresented groups face a complex set of barriers to achieving faculty status [@Coe2019]. For example, the dedication of women---who have children---to their work is perceived to be less than that of their colleagues, including men who also have children [@Firth1982; @Correll2007; @Fuegen2004].  Historically under-represented minorities in the United States (HURM), Asian/Asian-Americans, and women are all held to stricter competency standards and report having to work harder than Caucasian men to be perceived as legitimate scholars [@BlairLoy2017; @noauthor_seeking_2013]. Asian/Asian-Americans suffer from imposter syndrome at greater rates than other marginalized groups and Asian women report a lack of sponsors [@McGee2018; @wse_2016]. Increased performance expectations and patterns of exclusions are consistent themes in studies characterizing the HURM faculty experience [@pololi_race_2010; @hassouneh_experiences_2014].  While HURM and other marginalized groups share some experiences, differences including varying rates of hiring and tenure promotion mean that unique considerations are important for inclusion of each group [@Fang2000].

Here, we argue that to support the retention of faculty from marginalized groups at the professor level, universities should provide trainees with visual representation of themselves as successful scientists. Recent studies show that women in STEM benefit from women role models through improved belonging and self-efficacy [@Herrmann2016; @Drury2011]. Predictably, a lack of active inclusion also decreases self-efficacy in URMs, which can result in decreased feelings of belonging [@Lambert2020]. However, as the demographics of trainees have become more diverse, those who are not Caucasian men are lacking role models. Institutionalized racism and sexism [@Hardeman2018], defined as policies, societal norms, and ideologies that reinforce inequities, have played a large role in access to, inclusion in, and hiring policies at U.S. universities [@harvey_diversity_2011; @Stack2019]. Accordingly, faculty from marginalized groups are eliminated from the applicant pool and subsequent hires, leaving university policies and practices to be predominantly created by Caucasian men. Thus, institutionalized racism and sexism are perpetuated [@Iverson2007; @Tate2016]. Universities have begun adjusting hiring practices and creating initiatives to address inequitable hiring practices, but have had limited results [@Gibbs2016]. Considering that most faculties within the U.S. are still skewed towards Caucasian men, invited seminar series are a possible tool to provide marginalized trainees with representation of themselves as successful scientists.

Invited seminar series are common within biomedical departments across the United States [@nittrouer_gender_2018]. Usually, seminar series’ consist of faculty members selecting a scientist from another institution to visit their university and present their research, as well as meet with other faculty members and trainees. Named lectureships follow the same format but are decided by committee and are considered more prestigious because they are named in honor of prominent local scientists. These seminar series and lectureships provide an opportunity for trainees to be exposed to research outside of their department. Additionally, being an invited speaker provides the scientist with an opportunity to make future collaborations and build their own _curriculum vitae_ (CV). Scientists invited to give seminars are widely regarded as successful and the top in their field, providing an opportunity for trainees to be exposed to successful scientists in that field. Some studies have examined the addition of more women speakers at conferences to promote inclusion [@kalejta_gender_2017; @casadevall_presence_2014; @klein_speaking_2017],  however,  we have only identified one other study that has focused on invited speakers at universities [@nittrouer_gender_2018]. In their study, Nittrouer et al, examined 3,652 talks at 50 U.S. institutions in 2013 - 2014 and found that women faculty are less likely to be invited speakers [@nittrouer_gender_2018]. We have not been able to identify any publications examining scientific speaker diversity beyond gender or how department speaker series compares to trainee diversity in that department.

In this study, we examine and compare the demographics of invited speakers to Caucasian men in the Department of Microbiology and Immunology at the University of Michigan. Additionally, we compare invited-speaker demographics to the current trainee demographics to gauge if trainee demographics are being represented throughout the seminar series. Following our investigation, we proposed a policy change to the Department of Microbiology and Immunology in how invited speakers are selected to promote inclusion in our department and reduce unconscious bias. In order to facilitate inviting a more diverse group of scientists, we developed a set of resources that allow scientists, within the fields of microbiology and immunology, to self-identify any under-represented or under-served identity including: HURM, non-Caucasian/non-HURM (NCNH), or a Caucasian woman. These resources will promote inclusion and diversity by providing greater representation of all scientists and will provide hosts an opportunity to invite a more diverse group of scientists.

## Methods
```{r method, eval=TRUE, echo=FALSE, cache=FALSE,  warning=FALSE, message=FALSE, include=FALSE}
num_pd <- trainee_data %>% filter(Trainee_type == "postdoc") %>% nrow()

num_student <- trainee_data %>% filter(Trainee_type == "student") %>% nrow()

num_speaker <- nrow(speaker_data)
```

Each academic year, each faculty member in the Department of Microbiology and Immunology at the University of Michigan has the opportunity to invite one speaker per year for a weekly seminar series. Some of these seminar slots are dedicated to named lectureships, which are decided by committee, and three trainee-invited speakers. We analyzed the demographics of invited speakers and faculty hosts for five academic years (Fall 2014 - Spring 2019), and compared them to the current trainees when the data were analyzed (Spring 2019). Each speaker was only counted once and those listed as departmental faculty members or as a “host” at any point could not also be considered “invited speakers”. The list of faculty hosts was used as a proxy for faculty demographics since as hosts, these faculty members are visible representatives of the department. There were a total of `r num_speaker` invited speakers and paired hosts. The trainees were identified using departmental email lists that included masters students, doctoral students, and post-doctoral fellows. There were `r num_student` students and `r num_pd` post-doctoral fellows.

This is a retrospective study, thus speakers were not asked for their identities at the time of visit. Instead we hand-coded proxy demographics of the speakers, faculty hosts and trainees using first names, publicly available photos, and CVs [@Broderick2019; @Kimery2011; @Helmer2017; @Fox2019; @Gilbert1994]. Information from CVs, such as undergraduate institutions and activity in HURM groups helped inform our demographics by indicating identities that might be held by that individual. For instance, an undergraduate at a Puerto Rican university and activity in the non-profit Ciencia Puerto Rico suggests that even if the individual appears Caucasian, they probably identify as Puerto Rican and qualify as a URM. Because these data were collected from publicly available sources, this study was not submitted to an IRB for consideration. The presenting gender of each individual was inferred using a binary system (man/woman). Due to the low number of individuals in the study, race/ethnicity demographics were only split into three groups: Caucasian, Historically Under-represented Minority (HURM), and Non-Caucasian/Non-HURM (NCNH), each with a binary (yes/no) possibility. Caucasian was inferred using the current U.S. Census definition where those of Middle Eastern, European, and Russian descent are included. URM individuals include those of African-American, Indigenous American, Alaskan/Hawaiian Native, Latinx and/or Hispanic heritage (20 U.S. Code § 1067k), we use the HURM designation to recognize the history of enslavement and active oppression in the U.S.. The NCNH group predominantly included Asian/Asian-Americans, but also African immigrants [@Okonofua2013].

@@

trainee_summary <- tidy_trainee %>% 
  select(-Gender) %>% 
  filter(value == "y") %>% distinct_() %>% 
  group_by(demographic) %>% summarise(n = n()) %>% 
  mutate(prop = get_percent(n, sum(.$n)))
```

To understand the representation of women, we compared the proportion of women in each academic role. At the trainee level, more than half of students and postdoctoral fellows were women. That dropped to `r gender_combined[[1,4]]`% of faculty hosts and `r gender_combined[[3,4]]`% of the invited speakers (Fig. 1A). Of `r sum(lect_summary$n)` lectureships over the five year period, `r lect_summary[4,4]`% were awarded to women.

@@

This study found that the proportion of HURM and NCNH invited speakers were under-representative of the trainee populations for each group. Additionally, within the last 5 years, no HURM or NCNH woman was awarded a lectureship, despite that in 2017 non-Caucasians were 30% of the professoriate [@nsf_2019]. This means that the department is not providing non-Caucasian trainees with adequate representation of successful scientists and failing to support an inclusive environment in terms of visual faculty representation. We also found that the proportion of women as faculty hosts and speakers in our study population is equivalent to global estimates that 40% of microbiologists are women, though women only represent about 30% of academic biomedical faculty [@allagnat_gender_2017;@Jena2015]. Women are also over-represented as graduate students and postdoctoral fellows in this department. Overall, Caucasian scientists are over-represented as host faculty and invited speakers, compared to their presence as trainees, particularly when lectureships were considered.

We have not been able to identify any publications examining scientific speaker diversity beyond gender [@nittrouer_gender_2018]. This seems to be the first, which is concerning since conclusions drawn from gender-based studies are often framed, and considered, to be applicable to other marginalized groups (e.g., HURM). For instance, that African-American and Caucasian women benefit equally from the same policy change. This is a flawed assumption [@Armstrong2015]. While there is no doubt some overlap, each group remains marginalized due to a unique complex set of factors that cannot always be solved by gender-based solutions. The historical exclusion of HURMs by U.S. institutions means that they have a particular responsibility to improving the academic experience for these populations [@harvey_diversity_2011]. We therefore call on U.S. institutions to apply intersectional framing to their discussions and research.

Departments have different processes and criteria for selecting invited speakers, but it is a matter of pride to bring the best scientists possible. The barriers to achieving faculty status for HURMs, Caucasian women, and NCNH may also impact their speaking invitations. For instance, the perceived prioritization and commitments of women to family over work may cause faculty to doubt their acceptance of a speaking invitation, despite the prestigious nature of these invitations and evidence that men and women accept at similar rates [@nittrouer_gender_2018; @xu_gender_2008]. As a result, the faculty member may invite a different colleague who they feel is more likely to agree (and is a man). It may also be that the definition of “best” poses a problem to under-represented and under-served groups (e.g., women, HURM, and Asian) who are held to stricter competency standards [@BlairLoy2017;@noauthor_seeking_2013]. Some departments may only invite tenured faculty, which severely limits the number of potential speakers who are Caucasian women or non-Caucasian. Yet, another scenario is that pre-tenure faculty members invite prestigious, tenured faculty in their field to network and secure letters for their own tenure package. The increased burden of women and non-Caucasian scientists to prove competency decreases their likelihood to be considered for either tenure or as possible source of tenure letters. In particular, the proportion of HURM faculty at the Assistant and Associate Professor level is currently higher than at Full Professor so it will be difficult to increase speaker diversity if early-career researchers are not being considered [@whittaker_retention_2015]. Even when HURM speaker rates match the proportion of HURM faculty employment, HURM trainees will be represented at a significantly higher proportion. We argue that inclusion of marginalized faculty in seminar series is an important factor to increasing their representation among Associate and Full Professors. It is just one aspect of larger institutional change that is needed, but one that will benefit trainee experiences and the CVs of faculty from marginalized groups [@johnson_msphere_2019].

We recognize that our proxy demographics are a limitation of the analysis and want to acknowledge that biological sex (male/female) is not always equivalent to the gender that an individual presents as (man/woman), which is also distinct from the gender(s) that an individual self-identifies as. We also want to acknowledge that there are many other identities that are not captured in this limited analysis and that our personal implicit biases may have impacted our assignment of demographics. While our pilot study combined 5 years worth of seminars, our n is still quite low and we did not have other departments for comparison. Consequently, our results can not be generalized to other departments, fields, or universities. Another limitation to looking at a single department is that trainees may interact with faculty from many other departments, depending on their research and interests. Therefore, the individual experience of representation would vary by trainee. Future research needs to consider multiple departments, universities, and/or fields to bolster generalizability. There is a paucity of research on speaker identities other than their gender, so this also needs to be addressed in future studies. However, we caution that representation is a shallow metric on which a single-minded focus can cause more harm than help [@Iverson2007]. We recommend that future research also survey trainee and speaker experiences and trainee participation in seminar series to better understand the dynamics at play.

@@

We also used this opportunity to begin a conversation about the purpose of seminar speakers. Seminar speakers are sometimes invited to highlight their latest high-impact paper or to share the arc of discoveries they have made over several years of their career. While both are certainly worthwhile, there are other benefits to be gained from interacting with seminar speakers such as how to apply new techniques and how research is framed outside research-focused universities (Table 1). Thinking more broadly about what material is valuable during a seminar series may lead to more speakers from under-represented and under-served backgrounds, as well as more diversity of career paths. For some institutions, these suggestions represent more of a structural change to the departmental seminar series as speakers focusing on techniques or from non-research-intensive universities are usually invited as part of a professional development series. If these changes are to be implemented, many members of the department must agree to the value of including these seminars in the main departmental seminar series and these expectations must be clearly communicated by the leadership (Table 1). Departmental leadership can also ask individuals and groups to consider the unconscious biases that may be impacting their own speaker nominee lists to combat some of the barriers to inviting diverse speakers.

Presented with these ideas, several members of the Department of Microbiology and Immunology at the University of Michigan expressed interest in specific resources they might use to identify individuals from diverse identities, careers, and institutions. One suggestion is to use resources that are being organized by professional societies such as the American Society for Microbiology (ASM) and the Society for Advancing Chicanos/Hispanics & Native Americans in Science (SACNAS) (Table 1). We also chose to develop “Diversify” resources for the microbiology and immunology fields that provide a list of scientists from under-represented and under-served groups, that is not associated with a specific professional society. More information on the type of resources and how to establish a Diversify list is below. Using list resources like those available from professional societies as well as our “Diversify” resources are particularly useful as social science research has shown that the human brain is much better at recognizing and using information (such as a strong scientific speaker) from a list than it is at recalling the same information from memory [@Loftus1971; @Johnson2014].

We caution, however, that it is not enough to invite speakers from diverse identites`. An inclusive environment must be built within the department. Start by inviting all speakers to spend a few minutes describing their personal science journey and providing time for trainees to engage with the speaker. Trainee-speaker interactions can be encouraged by ensuring that trainees are the first to ask questions at the seminar’s conclusion and by scheduling a dedicated meeting time for trainees. As speaker schedules are being designed, departments should consider how they can foster an inclusive atmosphere during the speaker visits. Speakers should be asked prior to their visit if they have any dietary, movement, or other restrictions that should be accommodated during the visit. The identities of individuals the speaker is meeting with during their visit may also need to be considered; this is not to say that all identities of a speaker should be matched on their schedule or that a HURM speaker must have a meeting a HURM faculty member, but take care to ensure that speakers are meeting with faculty that reflect the diversity of thought and identity in the department. If a portion of the department is consistently not represented on speakers’ schedules (or extending invitations) this may reflect an opportunity for increased inclusivity in the department. Finally, speakers should be provided with ample opportunity to request meetings, not only with faculty but also with student groups or campus administrators who share similar interests. Through these steps departments can increase the diversity of speakers invited to their seminars while also increasing the impact of the seminar speakers.

@@

The website provides an interface to the Google forms and spreadsheets with template pages for viewing the list, adding a name to the list, and finding additional resources. Importantly, our website creation tool is hosted for free by GitHub, which provides a free website for each GitHub organization. Basic tools and skills required to set up a Diversify site include knowledge of, or experience with, the version control tool git, the web-tool GitHub, and a text editor. A tutorial in the DiversifyMicrobiology repository on GitHub provides links to these resources and instructions for adapting the tool to your own field (Table 1) [@Hagan2019_micro]. We caution creators of Diversify lists that the data voluntarily submitted to these lists is not eligible for study. IRB approval must be obtained prior to launching the list if that is a goal.

@@

To increase the retention of Caucasian women, HURM and NCNH trainees in the biomedical sciences, they must also be represented as experts. However, the invited speaker diversity at one department does not represent the diversity of trainees. To facilitate the identification and recruitment of individuals in these, and other, under-served groups, we have built a tool to create self-nominated, field-specific lists.

@@

## Author Contributions
A.K.H. collected the data, inferred demographics, analyzed the data, created the website, and wrote the methods and results. R.M.P. created the Google lists, forms, and website content and the description of their maintenance. J.L. wrote the introduction and provided conceptual advice. A.K.H. and J.L facilitated the policy change to the UM Department of Microbiology and Immunology. All authors contributed to preparing the final manuscript.

@@

\newpage
## References

@@

**Figure 1. The demographics of invited speakers, hosting faculty, and trainees.** A) The proportion of women in each academic role. B) The proportion of each academic role represented by individuals that are Caucasian (left), Historically Underrepresented Minorities (HURM, center) or Non-Caucasian/Non-HURM (NCNH, right). C-D) The percent of lectureships awarded to individuals that are C) Caucasian, HURM, or NCNH and D) Caucasian, HURM, or NCNH by gender.

**Table 1. Suggestions and resources to increase invited speaker diversity.**
